Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-14 paper.tex
Minor edits
rest of discussion and conclusion

%~ \subsection{Effect of Data Locality}
%~ % It should be important, in particular when there is contention and when
%~ % tasks are small. 
%~ We measure the effect of data locality by comparing the runs of Spark 
%~ on local disk (non-hatched green bars in Fig.~\ref{fig:results}) to the 
%~ ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
%~ local execution peaks at 3.2, reached at 750 chunks in experiment 3. \todo{This needs to be finished or removed.}

@@

@@

Writing to node-local storage in a cluster environment comes at a cost, for both
Nipype and Spark without HDFS.
% fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
%                  then the only solution is to recompute, which spark would do automatically.
%                  when nipype is run on lustre, it would avoid recomputing existing files, which Spark doesn't do.
%                   this niype feature is lost when not using lustre.

@@

% mention burst buffers or heterogeneous storage managers as potential solutions to write to local disk, 
% i.e., to benefit from data locality while getting your results on Lustre.

\subsection{Engine-Specific Comparisons}

@@

\section{Conclusion} % 1 page with refs
\label{sec:conclusion}
% When is it useful to use a Big Data engine?

>>>>

\subsection{Effect of Data Locality}
%~ % It should be important, in particular when there is contention and when
%~ % tasks are small. 
We measure the effect of data locality by comparing the runs of Spark
on local disk (non-hatched green bars in Fig.~\ref{fig:results}) to the 
ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
local execution peaks at 3.2, reached at 750 chunks in experiment 3. Overall, 
writing locally was usually preferrable to writing to Lustre, as a result of the
lower contention on local disk. Although it may be true, as stated in~\cite{Anantharayanan2011datalocality},
network bandwidths generally exceed that of disks, locality remains important 
as contention to a shared filesystem will be much higher than local disk.\todo{This needs to be finished or removed.}
The only time in our experiments where it was found that writing locally did not
have a huge impact over Lustre, was in experiments 1 and 4, at 1 iteration and 
when processing the MRI image, respectively. In both these scenarios, the Lustre 
writes did not impact performance as the data was able to be written to page cache
and flushed to Lustre asynchronously.

@@

With Nipype, MapNodes, which apply a given function to each element in a list, 
were found to be slower than the Node, which apply a function to a single element,
due to a blocking mechanism. For this reason, we selected to iterate through a 
series of Nodes in our code.

@@

Writing to node-local storage in a cluster environment comes at a cost, for both
Nipype and Spark without HDFS. When a node is lost, the node-local data is lost
with it. While Spark will recompute the lost partition automatically using lineage, 
Nipype will fail for all tasks requiring the data. Spark, however, will fail if 
RDDs of node-local filenames are shuffled, as the data associated with the filenames
will not be shuffled with the RDD

Nipype, when executed on Lustre, will checkpoint itself, ensuring resumption from
last checkpoint. This is particularly important in the case of compute-intensive 
applications, such as those found in neuroimaging. Spark also provides a checkpointing
mechanism, however, it requires HDFS.

It is common in neuroimaging applications for users to want access to all intermediate
data. Such a feature is currently only possible when writing to shared filesystem. 
It would also not be an option with Spark in-memory. To enable this burst buffers or 
heteorogeneous storage managers (e.g. Triple-H~\cite{tripleH}) could be used to ensure
fast processing and that all outputs (including intermediate outputs) will be 
sent asynchronously to the shared filesystem. \todo{double check triple-H}

% fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
%                  then the only solution is to recompute, which spark would do automatically.
%                  when nipype is run on lustre, it would avoid recomputing existing files, which Spark doesn't do.
%                   this niype feature is lost when not using lustre.

@@

% mention burst buffers or heterogeneous storage managers as potential solutions to write to local disk, 
% i.e., to benefit from data locality while getting your results on Lustre. 
\subsection{Engine-Specific Comparisons}

It was expected that Spark would experience longer processing times, particularly
with the small datasets, due to Java serializing. This was not found to be the case. 
Unlike Spark, Nipype performs a more thorough provenance capture, potentially owing to longer
processing times.

@@

\section{Conclusion} % 1 page with refs
\label{sec:conclusion}

Big Data performance optimization strategies help improve performance of typical 
neuroimaging applications. Our experiments indicate that overall, in-memory computing 
enables greater speedups than what can be obtained by using page cache and tmpfs. 
While page cache and tmpfs do give memory-like performance, they are likely to fill
up faster than memory, leading to increased performance penalties when compared
to in-memory computing.

Data-locality was also found to play an important role in 
application performance. Local disk was found to perform between than the shared
filesystem, despite having lower bandwidth, due to increased contention on the 
shared filesystem. As local disk will generally have less contention than a 
shared filesystem. It is recommended to have the data stored locally. Although, 
storing data locally may come at a cost to fault tolerance.

Although a more thorough analysis of lazy evaluation remains to be performed, it
is speculated that this may be the cause of the general performance difference 
between Spark and Nipype. Furthermore, it is found that lazy evaluation optimizations increase 
the likelihood of cache hits, thus improving overall performance.

Even though Big Data strategies are beneficial to the processing of Big 
neuroimaging data, it is estimated that it would require running a functional MRI
dataset with 200 concurrent subjects for any noticeable impact using our Lustre 
bandwidth estimate.

To further improve neuroimaging data processing times, scheduling strategies for 
Big Data application in a multitenant environment need to be investigated. 
Additionally, workflow-aware cache eviction policies that maximizes page cache 
usage for the workflows should be investigated.
% When is it useful to use a Big Data engine?
